Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{\textbf{Ok .....................!!!!!!!!!!!!!!!!!!..................}}

### 2015-09-05 paper.tex
Finished the general review.

% first the title is needed
\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{Ok}}
% a short form should be given in case it is too long for the running head
\titlerunning{ATP Based on the PCF Calculus}

@@

Let $X = \{x_1,\ldots,x_k\}$ be a set of variables, $A = \{A_1,\ldots,A_m\}$ be a set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be a set of FOFs. Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \& \ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F^{\vee}$, where $F^{\vee}$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F^{\&}$, where $F^{\&}$ means conjunctions of all formulas from $F$.

@@

Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics. But type quantifiers are stable in languages of another applied fields.

@@

\begin{definition}[Splitting]
\label{splitting}
Let $B = \exists_XA\colon\Psi$, and $Q = \forall_YD\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots,\exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(B,Q) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator. We say that $B$ is split by $Q$. Obviously, $split(B,\forall_YD) = split(B,\forall_YD\colon\emptyset) = \emptyset$.
\end{definition}

@@

%=======================================================================
%==========================EXPERIMENTS==================================
%=======================================================================
\section{Experiments}
Testing the developed prover is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org). This library is the de facto standard for testing ATP provers. The type of problems that were chosen are the problems of FOF without equality. Total number of problems is 1221 with rating from 0.0 (very easy) to 1.0 (unresolved). Total number of solved problems is 780. The maximal rating of solved problem is 0.62. The best results among domains presented in the Table~1.

\begin{table}
\caption{Domains}
\begin{tabular}{|p{0.3\linewidth}|p{0.4\linewidth}|p{0.1\linewidth}|}

@@

10 hardest problems solved by our prover is presented in the Table~2.
\begin{table}
\caption{Hardest problems solved by our prover}
\begin{tabular}{|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|}

@@

Prover shown good results for problems, whose formalization in the PCF language contains large conjuncts (type conditions).

@@

The abstract syntax tree is passed to translation component that imports the tree, translates the predicate formula or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language, performs additional reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.

\subsection{Example of Formula}
Problem SYN380+1.p from TPTP library in our prover input language has a following form:
\begin{quote}
\texttt{\raggedright\noindent
\{\\
~~e[][]~\{\\
~~~~a[W][big\_r(W,W)]~\{\\
~~~~~~e[][False]~\{\}\};\\
~~~~a[X,Y][]~\{\\
~~~~~~e[][big\_r(X,Y)]~\{\};\\
~~~~~~e[Z][big\_q(Y,X)]~\{\\
~~~~~~~~a[][big\_q(Z,Z)]~\{\\
~~~~~~~~~~e[][False]~\{\}\}\}\}\}\\
\}\\
}
\end{quote}

@@

The results of solving this problem by our prover and other provers are presented in Table~3.

\begin{table}
\caption{COM003+1.p solving}
\begin{tabular}{|p{0.2\linewidth}|p{0.35\linewidth}|}

@@

%=============================
\subsubsection{Lazy concretization.}
Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum. The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy. The decrease of the number of steps leads to a decrease of the problem solving time. Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching.

%=============================
\subsubsection{$k,m$--condition.}
The efficiency of this strategy reveals in the problems with disjunctive branching. In all cases of acceptable application of this strategy and correctly chosen parameters, the noticeable improvement is achieved.
%===========================
\subsubsection{Parallel stratgies.}
The important quality of parallel schemes of algorithms is their scalability, i.e., the performance improvement respect to increase of the number of computational nodes (CNs). That’s why the main characteristic of the prover productivity is not the time of execution of the program, but the ratio of program execution time in parallel mode on a given number of CNs to the program execution time on a single CN. The ratio is accounted for various number of CNs.

Experiments were carried out on the logical problems which PCF--formalizations have necessary properties for the parallel strategies: disjunctive branching, large number of questions, large conjuncts in questions. In Fig.~\ref{fig:parallel} the results of experiments are shown.

>>>>

% first the title is needed
\title{Software for Automated Theorem Proving Based on the Calculus of Positively Constructed Formulas\thanks{\textbf{Ok .....................!!!!!!!!!!!!!!!!!!..................}}}
% a short form should be given in case it is too long for the running head
\titlerunning{ATP Based on the PCF Calculus}

@@

Let $X = \{x_1,\ldots,x_k\}$ be a set of variables, $A = \{A_1,\ldots,A_m\}$ be a set of atomic formulas, and $F = \{F_1,\ldots,F_n\}$ be a set of FOFs. Then the following formulas $((\forall x_1) \ldots (\forall x_k) (A_1 \& \ldots \& A_m \rightarrow (F_1 \vee \ldots \vee F_n)))$ and $((\exists x_1) \ldots (\exists x_k) (A_1 \&$ $\ldots \& A_m \& (F_1 \& \ldots \& F_n)))$ are denoted as  $\forall_XA\colon F$ and $\exists_XA\colon F$ respectively, keeping in mind that the $\forall$--quantifier corresponds to $\rightarrow F^{\vee}$, where $F^{\vee}$ means disjunction of all formulas from $F$, and $\exists$--quantifier corresponds to $\& F^{\&}$, where $F^{\&}$ means conjunctions of all formulas from $F$.

@@

\textbf{Originally, the term ``type quantifier'' was introduced by N.~Bourbaki \cite{Bourbaki} as part of notation for formalization of mathematics. But type quantifiers are stable in languages of another applied fields.}

@@

\begin{definition}[Splitting]
\label{splitting}
Let $B = \exists_XA\colon\Psi$, and $Q = \forall_YD\colon\Upsilon$, where $\Upsilon = \{\exists_{Z_1}C_1\colon\Gamma_1,\ldots, \exists_{Z_n}C_n\colon\Gamma_n\}$ then $split(B,Q) = \{\exists_{X \cup {Z_1}'} A \cup {C_1}'\colon\Psi \cup {\Gamma_1}',\ldots,$ $\exists_{X \cup {Z_n}'} A \cup {C_n}'\colon\Psi \cup {\Gamma_n}'\}$, where $'$ is a variable renaming operator. We say that $B$ is split by $Q$. Obviously, $split(B,\forall_YD) = split(B,\forall_YD\colon\emptyset) = \emptyset$.
\end{definition}

@@

%=======================================================================
%==========================EXPERIMENTS==================================
%=======================================================================
\section{Experiments}
Testing the prover under development is being performed on problems form TPTP \cite{tptp} library version 5.4.0 (Thousands of Problems for Theorem Provers, www.tptp.org). This library is the de facto standard for testing ATP provers. The type of problems that were chosen are the problems of FOF without equality. Total number of problems is 1221 with rating from 0.0 (very easy) to 1.0 (unresolved). Total number of solved problems is 780. The maximal rating of solved problem is 0.62. The best results among domains presented in Table~\ref{tbl:1}.

\begin{table}
\caption{Domains}\label{tbl:1}
\begin{tabular}{|p{0.3\linewidth}|p{0.4\linewidth}|p{0.1\linewidth}|}

@@

Ten hardest problems solved by our prover is presented in Table~\ref{tbl:2}.
\begin{table}
\caption{Hardest problems solved by our prover}\label{tbl:2}
\begin{tabular}{|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|p{0.2\linewidth}|}

@@

Prover showed good performace for problems, whose formalization in the PCF language contains large conjuncts (type conditions).

@@

The abstract syntax tree is passed to translation component that imports the tree, translates the FOF or CNF, performs some reduction of subformulas by applying a set of rules, converts the formulas into the PCF language. Translator also performs additional trivial reduction of the PCFs, renames the quantifier variables, and, at the last step, outputs the conversion result as textual representation of the problem in the language of the prover.

\subsection{Example of formula}
Problem SYN380+1.p from TPTP library in our prover input language has a following form:
\begin{quote}
\texttt{\raggedright\noindent
\{\\
~~e[][]~\{\\
~~~~a[W][big\_r(W,W)]~\{\\
~~~~~~e[][False]~\{\}\};\\
~~~~a[X,Y][]~\{\\
~~~~~~e[][big\_r(X,Y)]~\{\};\\
~~~~~~e[Z][big\_q(Y,X)]~\{\\
~~~~~~~~a[][big\_q(Z,Z)]~\{\\
~~~~~~~~~~e[][False]~\{\}\}\}\}\}\\
\}\\
}
\end{quote}

@@

The results of solving this problem by our prover and other provers are presented in Table~\ref{tab:3}.

\begin{table}
\caption{Solution timing for COM003+1.p}\label{tab:3}
\begin{tabular}{|p{0.2\linewidth}|p{0.35\linewidth}|}

@@

In this section we list some points of the implemented strategies practical usage experience.

%=============================
\subsubsection{Lazy concretization.}
Lazy concretization strategy allows one in most cases to find a substitution for an unconfined variable in acceptable time as compared to the sequential enumeration of Herbrand universum. The constraints are the most valuable part of this strategy, which allow to decrease the number of steps of logical inference in comparison to the unconstrained strategy. The decrease of the number of steps leads to a decrease of the problem solving time. \textbf{Application of the constraint~1 is the best way of the inference time reduction in the case when a formula has questions with disjunctive branching and the constraint~2 in case if the formula contains no questions without disjunctive branching}.

%=============================
\subsubsection{$\boldsymbol{k,m}$--constraint.}
The value of this strategy reveals in the problems with disjunctive branching. A noticeable improvement of efficiency is achieved in all cases of acceptable application of this strategy and correctly chosen parameters.

%===========================
\subsubsection{Parallel stratgies.}
The important quality of parallel schemes of algorithms is their scalability, i.e., the performance improvement respect to increase of the number of computational nodes (CNs). That’s why the main characteristic of the prover productivity is not the time of execution of the program, but the ratio of program execution time in parallel mode on a given number of CNs to the program execution time on a single CN. The ratio has been evaluated for various number of CNs.

Experiments were carried out on the logical problems which PCF--for\-ma\-li\-za\-ti\-ons having necessary properties for the parallel strategies utilization: disjunctive branching, large number of questions, large conjuncts in questions. In Fig.~\ref{fig:parallel} the results of experiments are shown. \textbf{In the figure ``Strategy 1'' denotes ... and ``Strategy 2'' denotes ...}
